The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2013-03-13 script/en/chapters/cryptomethods.tex
CT-Script en: Minor change (SAT solver added in chap 1).

% HACK to fix warning "destination with the same identifier .. has already been used, ...":
\makeatletter \renewcommand{\thepage}{~\csname @arabic\endcsname \c@page~} \makeatother
\hypertarget{Kapitel_1}{}
\chapter{Security Definitions and Encryption Procedures}
\label{Label_Kapitel_1}
(Bernhard Esslinger, Joerg-Cornelius Schneider, May 1999; Updates Dec. 2001, Feb. 2003,
June 2005, July 2007, Jan. 2010, Dec. 2012)

@@

@@

\bibitem[Mansoori-etal2007]{cm:Mansoori-etal2007} \index{Mansoori, Bizaki 2007} 
       S. Davod. Mansoori, H. Khaleghei Bizaki, \\
       {\em On the vulnerability of Simplified AES Algorithm Against Linear Cryptanalysis}, \\
       In: IJCSNS International Journal of Computer Science and Network Security, Vol.7 No.7,
           July 2007, pp~257-263 \\
       See also:\\
       \url{http://paper.ijcsns.org/07_book/200707/20070735.pdf}

>>>>

% HACK to fix warning "destination with the same identifier .. has already been used, ...":
\makeatletter \renewcommand{\thepage}{~\csname @arabic\endcsname \c@page~} \makeatother
\hypertarget{Kapitel_1}{}
\chapter{Security Definitions and Encryption Procedures}
\label{Label_Kapitel_1}
(Bernhard Esslinger, Joerg-Cornelius Schneider, May 1999; Updates Dec. 2001, Feb. 2003, June 2005, July 2007, Jan. 2010, Mar. 2013)

@@

% --------------------------------------------------------------------------
\subsection{Algebraic or algorithmic cryptanalysis on symmetric algorithms}
\index{Attack!algebraic}
\index{SAT solver}
\label{Algebraic-gegen-Symmetr}

There are different modern methods attacking the structure of a problem directly or after a transformation of the problem. One of the attack methods is based on the satisfiability problem (SAT)%
\footnote{%
 \url{http://en.wikipedia.org/wiki/Boolean_satisfiability_problem}
}.

%\vskip +25 pt
\paragraph*{Description of a SAT solver}\index{SAT solver}\mbox{}
\hypertarget{ht_SAT-Solver}{}

An old and well-studied problem in computer science is called the SAT problem. Here, for a given Boolean formula, it's the task to find out whether there is an assignment of the variables, so that the evaluation result of the formula is 1.

Example: The Boolean formula "A AND B" evaluates to 1, if and only if A=B=1. For the formula "A AND NOT(A)" there exists no assignment of its variable A, so that the formula is evaluated to the value 1.

For larger Boolean formulas, it is not easy to determine if an assignment exists for which the formula can be evaluated to 1 (this problem belongs to the NP-complete problems). Therefore specific tools have been developed to solve this problem for general Boolean formulas, so called SAT solvers\footnote{
    With {\bf CT2}\index{CT2}\index{CrypTool 2} you can execute
    a SAT solver -- using in the Startcenter
    {\bf Templates
    \textbackslash{} Mathematics \textbackslash{}
    SAT Solver (Text Input)}  and  {\bf SAT Solver (File Input)}.
    }. As has been found, SAT solvers can also be used to attack cryptographic systems.

%\vskip +25 pt
\paragraph*{SAT solver based cryptanalysis}\index{SAT solver}\mbox{}
\hypertarget{ht_SAT-Solver_Cryptanalysis}{}

The general approach to use SAT solvers in cryptanalysis is very straightforward: First, the cryptographic problem, e.g. finding the symmetric key or an inversion of a hash function, is translated into a SAT problem. This is done in a way that the solution of the SAT problem solves the original cryptographic problem. Therefore, the SAT solver can be used to find a solution to the SAT problem. The paper by Massacci \cite{cm: Massacci2000} describes the first known usage of a SAT solver in this context. Unfortunately, very soon it turned out that such a general approach cannot be used efficiently in practice. This is due to the fact that the cryptographic SAT problems are very complex and the runtime of a SAT solver increases exponentially with the problem size. Therefore, in modern approaches SAT solvers are used only for solving partial problems of cryptanalysis. A good example for this is described in the paper by Mironov and Zhang \cite{cm: Mironov2006}. They demonstrate the usage of a SAT solver in an attack on hash functions, where the SAT solver is used to solve some partial problems in a very efficient way.

@@

\bibitem[Mansoori-etal2007]{cm:Mansoori-etal2007} \index{Mansoori, Bizaki 2007} 
       S. Davod. Mansoori, H. Khaleghei Bizaki, \\
       {\em On the vulnerability of Simplified AES Algorithm Against Linear
       Cryptanalysis}, \\
       In: IJCSNS International Journal of Computer Science and Network Security,
       Vol. 7 No. 7, July 2007, pp~257-263 \\
       See also:\\
       \url{http://paper.ijcsns.org/07_book/200707/20070735.pdf}

\bibitem[Massacci2000]{cm: Massacci2000} \index{Massacci 2000}
       Fabio Massacci, Laura Marraro \\
       {\em Logical cryptanalysis as a SAT problem: Encoding and analysis},\\ 
       In: Journal of Automated Reasoning Security,
       Vol. 24, 2000, pp~165-203.

### 2013-03-13 script/en/chapters/numbertheory.tex
CT-Script en: Minor change (SAT solver added in chap 1).

\noindent If you don't want to use factorization but ask the question like in classic
encryption, you may want to find out:
How many possible private keys $(n, d)$ are there for a given key size range
$n \in [a, b]$?\footnote{%
See chapter~\ref{L_nt_Num-of-d-mod-26} (``\nameref{L_nt_Num-of-d-mod-26}''),
p.~\pageref{L_nt_Num-of-d-mod-26}, which deals with the special case $n=26$.
}
%% (this means, only a part of each public key $(n, e)$ is known).

\noindent Sage source code~\ref{nt_sagesample_Count_RSA_Keys} below defining the
function \verb#count_Number_of_RSA_Keys# can answer this question concretely
(if the modulus is not too big).\footnote{%
%\begin{compactitem}
%       \item
 a) Calling \verb#sage: count_Number_of_RSA_Keys(100, 1000)# means to consider the interval
$[100, 1000]$ for $n$. $n$ is defined by the two primes $p, q: n = p*q$.
So the primes can have the maximal value $500$ because $2 * 500 =1000$.\\
The number of possible combinations of primes is $comb = 258$.\\
The number of primes in the given range is $143$.\\
The number of private keys is $34,816$.

\noindent %       \item
 b) Calling \verb#sage: count_Number_of_RSA_Keys(100, 100, True)# has the
following output:\\
   - Number of private keys for modulus in a given range: 0\\
   - Number of primes in a given range: 0\\
   The reason for that is, that with this call only $n=100$ is considered, and
   die Funktion nur semiprime $n$ untersucht: $100$ is not semi prime\index{Prime number!semi prime}\index{Prime number!half prime}\index{Half prime},
   this means $100$ is not the product of only two primes.
%\end{compactitem}
}

>>>>

\noindent If you don't want to use factorization but ask the question like in classic
encryption, you may want to find out:
How many possible private keys $(n, d)$ are there for a given key size range
$n \in [a, b]$?\footnote{%
Chapter~\ref{L_nt_Num-of-d-mod-26} (``\nameref{L_nt_Num-of-d-mod-26}''),
p.~\pageref{L_nt_Num-of-d-mod-26} deals with the special case $n=26$.}
%% (this means, only a part of each public key $(n, e)$ is known).

\noindent Sage source code~\ref{nt_sagesample_Count_RSA_Keys} below defining the
function \verb#count_Number_of_RSA_Keys# can answer this question concretely
(if the modulus is not too big).\footnote{%
%\begin{compactitem}
\newlength{\saveleftmargini}
\setlength{\saveleftmargini}{\leftmargini}
\setlength{\leftmargini}{0em}% for example to outdent verse
\settowidth{\versewidth}{xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx xxxxx}
\vspace{-\baselineskip} % TODO: Geht das nicht anders?
\begin{verse}[\versewidth]
 a) Calling \verb#sage: count_Number_of_RSA_Keys(100, 1000)# means to consider the interval
$[100, 1000]$ for $n$.
\verselinebreak $n$ is defined by the two primes $p, q: n = p*q$.
\verselinebreak So here one prime can have the maximal value $500$ because $2 * 500 =1000$
(while then the other prime will have the smallest possible prime value $2$).\\
\vin The number of possible combinations of primes is $comb = 258$.\\
\vin The number of primes in the given range is $143$.\\
\vin The number of private keys is $34,816$.
\end{verse}
%\noindent %       \item
\begin{verse}[\versewidth]
 b) Calling \verb#sage: count_Number_of_RSA_Keys(100, 100, True)# has the
following output:\\
\vin    - Number of private keys for modulus in a given range: 0\\
\vin    - Number of primes in a given range: 0\\
\vin    The reason for that is, that with this call only $n=100$ is considered, and the
   function investigates only\\
\vin semiprime $n$: $100$ is not semi prime\index{Prime number!semi prime}\index{Prime number!half prime}\index{Half prime},
   this means $100$ is not the product of only two primes.
\end{verse}
\setlength{\leftmargini}{\saveleftmargini}% restore original value%\end{compactitem}
}

### 2013-03-13 script/en/script-en.tex
CT script: Minor formal changes in chapter 4
Several minor changes (for consistency and correctness), mostly in chap 4.
Added chap 8 (HE) from Dr. Martin Franz. Plus some minor changes.
center citations
CT Script: Added 48th M-Prime and some minor changes (DE and EN)
CT-Script en: Minor change (SAT solver added in chap 1).

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{fancyvrb}
\usepackage{makeidx}
\usepackage[pdftex]{graphicx}
\usepackage{pdfpages}   % SR_2009: added new cover

@@

\floatstyle{boxed}
\newfloat{ctsquote}{H}{loq}   % ctsQuotes-from-Books, List-of-Quotes
\floatname{ctsquote}{Quote}

\floatstyle{ruled}

@@

% \includepdf[page=1]{figures/coverv1_en-300dpi.pdf}
% \includepdf[page=1]{figures/coverv1_en.pdf}
\includepdf[page=1]{figures/coverv1_en_draft.pdf}

@@

\input{chapters/cryptomethods.tex}
\input{chapters/paper_and_pencil.tex}
\input{chapters/primes.tex}
\input{chapters/numbertheory.tex}
\input{chapters/moderncryptography.tex}
\input{chapters/digitalsignatures.tex}
\input{chapters/ellipticcurves.tex}
\bibliographyunit[\chapter]
\bibliography*{../de/references}
\bibliographystyle*{plain}
\input{chapters/crypto2020.tex}
\bibliographyunit

@@

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Quotes}
\listof{ctsquote}{List of Quotes}
\index{Quotes}
\label{sc:List-of-Quotes}

@@

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{fancyvrb}
\usepackage{makeidx}
\usepackage[pdftex]{graphicx}
\usepackage{pdfpages} %SR_2009: added new cover
%\usepackage{enumitem} %BE_2012: eliminate space after end{itemize} didn't work
\usepackage{paralist} %BE_2012: eliminate space after end{itemize} worked
                      %         with compactitem
\usepackage{arydshln} %BE_Jan2013: To use hdashline
\usepackage{verse}    %BE_Feb2013: To use \settowidth{\versewidth}{xxx}

@@

\floatstyle{boxed}
\newfloat{ctsquotefloat}{H}{loq}   % ctsQuotes-from-Books, List-of-Quotes
\floatname{ctsquotefloat}{Quote}   % ctsQuotes with centering
\newenvironment{ctsquote}{\begin{ctsquotefloat}\center}{\end{ctsquotefloat}}
\floatstyle{ruled}

@@

% \includepdf[pages=1]{figures/coverv1_en-300dpi.pdf}
% \includepdf[pages=1]{figures/coverv1_en.pdf}
\includepdf[pages=1]{figures/coverv1_en_draft.pdf}

@@

\input{chapters/cryptomethods.tex}
\input{chapters/paper_and_pencil.tex}
\input{chapters/primes.tex}
\input{chapters/numbertheory.tex}
\input{chapters/moderncryptography.tex}
\input{chapters/digitalsignatures.tex}
\input{chapters/ellipticcurves.tex}
\input{chapters/homomorphicencryption.tex}
\bibliographyunit[\chapter]
\bibliography*{../de/references}
\bibliographystyle*{plain}
\input{chapters/crypto2020.tex}
\bibliographyunit

@@

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Quotes}
\listof{ctsquotefloat}{List of Quotes}
\index{Quotes}
\label{sc:List-of-Quotes}

@@

\newpage
\newpage
